Corotating Clumps Around Adolescent Low-Mass Stars: Four Years of Complex Quasiperiodic Variables from TESS

### 2023-08-13 paper/ms.tex
update ms

Most CQVs in our catalog did not show infrared excesses in the W1-W4
bands, which is typical for this class of object
\citep{2017AJ....153..152S}.  Visually inspecting the SEDs of our
\nallcands\ star sample, we labeled two objects as having ``good''
infrared excesses (both W3 and W4 at $>$3$\sigma$ above the
photospheric prediction), and three as ``possible'' infrared excess
(only W3 at $>$3$\sigma$).

The two ``good'' IR excesses belonged to TIC~193136669 (TWA~34) and
TIC~57830249 (TWA~33).  Both are in the TW Hydrae association
($\approx$10\,Myr), and have relatively long periods of 38\,hr and
44\,hr respectively.  In our initial labeling, we labeled both as
``ambiguous'' CQVs because the dips in their light curves did not show
the rigid periodicity typical of CQVs;  their periods were also
relatively long.  However, inspection of further sectors clarified
that both sources are dippers (see online plots in
Figure~\ref{fig:vet}).  Independently, TIC 193136669 has a cold
molecular disk based on observed 1.3\,mm continuum emission and
resolved Keplerian $^{12}{\rm CO}(2-1)$ emission
\citep{2015A&A...582L...5R}.  It was also labeled a dipper by
\citet{2022ApJS..263...14C}; we agree with their designation, and
label it an ``impostor'' CQV in Table~\ref{tab:thetable}.  TIC
57830249 (TWA~33) also has previously detected 1.3\,mm continuum
emission \citep{2015A&A...582L...5R}, suggestive of cold dust grains
being present.  It is also a dipper.  Section~\ref{subsec:discdippers}
highlights plausible evolutionary connections between CQVs and dippers
in light of these ``misclassifications'' .

Our three ``possible'' infrared excesses were TIC~405910546,
289840926, and 244161191.  After a literature search, we concluded
that none have clear evidence for the presence of a disk.
TIC~405910546, in LCC, shows a unique TESS light curve, reminiscent of
a $P$=38\,hr singly-eclipsing binary, except with additional
substructure during each eclipse that resembles the CQV phenomenon
more than any other variability of which we are aware.  TIC~289840926
($\beta$~Pic moving group, $P$=4.8\,hr), shows what we believe is a
clear CQV signal, but has no definitive evidence for a large, dusty
disk.  TIC~244161191 (TOI-278), in Columba, also has no definitive
evidence for a large disk.  It is however ``multi-periodic''---in
addition to the 7.17\,hr CQV signal, this source shows a superposed
8.39\,hr sinusoidal signal, probably from an unresolved neighboring
star or binary companion.

@@

\begin{figure*}[!tp]
	\begin{center}
		\centering
		\includegraphics[width=0.98\textwidth]{f5.pdf}
		\vspace{-0.35cm}
		\caption{
			{\bf LP 12-502 (TIC~402980664) light curve}, where each time
			chunk represents one TESS orbit.  Data were acquired in Sectors
			18-19, 25-26, 53, and 57-58.  Flares are drawn in gray.  The red
			vertical lines highlight apparently instantaneous state changes
			in the shape of the dip pattern.  The light curve is binned to
			15-minute intervals so that there are 96 points per day, and
			each point is connected by a line.  Data gaps longer than
			15-minutes are not interpolated; if data are missing, nothing is
			plotted.
		}
		\label{fig:lplc}
	\end{center}
\end{figure*}

@@

We can estimate the mass of a transiting cloud by first converting the
transit depth to an effective cloud radius, $R_{\rm cloud}$.  For most
CQVs in Figure~\ref{fig:cqvs}, this typically yields
$\approx$2-20\,$R_\oplus$.  This size can be converted into a mass
estimate by requiring the optical depth $\tau$ of the cloud to be at
least unity and by positing some composition for the cloud.  In other
words, a minimum constraint on the number density follows by requiring
the cloud to be optically thick.  For cases like LP 12-502, this is
reasonable because the transit duration implies $R_{\rm cloud}\ll
R_\star$.  Carrying out the relevant calculation assuming the
occulting material is dust grains 1\,$\mu$m in size,
\citet{2023MNRAS.518.4734S} find minimum cloud masses of order
$10^{12}$\,kg (their Eq.~23), which scale linearly with both the
optical depth and dust grain radius.  This is comparable to a small
asteroid; the asteroid belt itself has a mass of order
$\approx$10$^{21}$\,kg \citep{2019Icar..319..812P}.  The previously
discussed cool gas prominences would need to have masses of order
$10^{14}$\,kg \citep{1990MNRAS.247..415C}, about 100$\times$ larger
than the requisite dust mass.

Our observations provide a direct measurement of how often dips appear
and disappear, both due to sudden state-switches, and due to more
gradual, secular evolution.  For instance, LP~12-502 showed three
``state-switch'' events over the six months of available TESS
observations, during cycles 248-264, 299-315, and 1233-1247.  In each
case, a dip ``turned off''.  It is plausible to imagine that these
events correspond to either mass being ejected from corotation, or
perhaps being accreted onto the star.  In either case, the
corresponding $\dot{M} \equiv M\cdot {\rm d}N/{\rm d}t$ time-averaged
over six months is $\approx$3$\times$$10^{-18} M_\odot\,{\rm yr}^{-1}
\approx 1\times10^{-12} M_\oplus\,{\rm yr}^{-1}$.  Considered
cumulatively over the $\approx$$10^8$ years for which the CQV
phenomenon is observed, this yields a cumulative moved dust mass of
$10^{-4}\,M_\oplus$, of order the Solar System's asteroid belt.  If
the occulting material is gas, the masses involved would be of order
100 times larger.  For cases in which we observe the {\it growth} of
dips, such as the Sector~29 data for TIC~224283342, or Sector~5 of
TIC~294328885, the dip depths typically increase by of order a few
percent over ten to twenty days.  This growth rate yields a mass flux
one order of magnitude larger than the earlier estimate.

@@

An argument against the connection between CQVs and the $\sigma$~Ori~E
analogs is that the light curve of $\sigma$~Ori~E is much simpler than
those in Figure~\ref{fig:cqvs}, with only two broad local minima, and
one ``hump'' \citep{2005ApJ...630L..81T, 2022ApJ...924L..10J}.  Within
the model proposed by Townsend et al., this is explained by the
simplicity of the star's magnetic field, which to first-order is a
dipole, similar to most other magnetic B stars
\citep{2007A&A...475.1053A,2009ARA&A..47..333D}.  This approximation
matches the qualitative behavior of both the broadband light curves,
and the line-profile variations seen in hydrogen, helium, and carbon
\citep{2012MNRAS.419..959O}.

@@

\clearpage

>>>>

Most CQVs in our catalog did not show infrared excesses in the W1-W4
bands, which is typical for this class of object
\citep{2017AJ....153..152S}.  Visually inspecting the SEDs of our
\nallcands\ star sample and the WISE images available through IRSA, we
labeled two objects as having reliable infrared excesses (both W3 and
W4 at $>$3$\sigma$ above the photospheric prediction).

The two IR excess sources were TIC~193136669 (TWA~34) and TIC~57830249
(TWA~33).  Both are in the TW Hydrae association ($\approx$10\,Myr),
and have relatively long periods of 38\,hr and 44\,hr respectively.
In our initial labeling, we labeled both as ``ambiguous'' CQVs because
the dips in their light curves did not show the rigid periodicity
typical of CQVs;  their periods were also relatively long.  However,
inspection of further sectors clarified that both sources are dippers
(see online plots in Figure~\ref{fig:vet}).  Independently, TIC
193136669 has a cold molecular disk based on observed 1.3\,mm
continuum emission and resolved Keplerian $^{12}{\rm CO}(2-1)$
emission \citep{2015A&A...582L...5R}.  It was also labeled a dipper by
\citet{2022ApJS..263...14C}; we agree with their designation, and
label it an ``impostor'' CQV in Table~\ref{tab:thetable}.  TIC
57830249 (TWA~33) also has previously detected 1.3\,mm continuum
emission \citep{2015A&A...582L...5R}, suggestive of cold dust grains
being present.  It is also a dipper.  Section~\ref{subsec:discdippers}
highlights plausible evolutionary connections between CQVs and dippers
in light of these ``misclassifications'' .

% Our three ``possible'' infrared excesses were TIC~405910546,
% 289840926, and 244161191.  After a literature search, we concluded
% that none have clear evidence for the presence of a disk.
% TIC~405910546, in LCC, shows a unique TESS light curve, reminiscent of
% a $P$=38\,hr singly-eclipsing binary, except with additional
% substructure during each eclipse that resembles the CQV phenomenon
% more than any other variability of which we are aware.  TIC~289840926
% ($\beta$~Pic moving group, $P$=4.8\,hr), shows what we believe is a
% clear CQV signal, but has no definitive evidence for a large, dusty
% disk.  TIC~244161191 (TOI-278), in Columba, also has no definitive
% evidence for a large disk.  It is however ``multi-periodic''---in
% addition to the 7.17\,hr CQV signal, this source shows a superposed
% 8.39\,hr sinusoidal signal, probably from an unresolved neighboring
% star or binary companion.

@@

\begin{figure*}[!tp]
	\begin{center}
		\centering
		\includegraphics[width=0.98\textwidth]{f5.pdf}
		\vspace{-0.3cm}
		\caption{
			{\bf LP 12-502 (TIC~402980664) light curve}, where each time
			chunk represents one TESS orbit.  Data were acquired in Sectors
			18-19, 25-26, 53, and 57-58.  Flares are drawn in gray.  
			The light curve is binned to
			15-minute intervals so that there are 96 points per day, and
			each point is connected by a line.  Data gaps longer are not interpolated; if data are missing, nothing is
			plotted.
			The red
			vertical lines highlight apparently instantaneous state changes
			in the shape of the dip pattern.  
		}
		\label{fig:lplc}
	\end{center}
\end{figure*}

@@

We can estimate the mass of a transiting cloud by first converting the
transit depth to an effective cloud radius, $R_{\rm cloud}$.  For most
CQVs in Figure~\ref{fig:cqvs}, this typically yields
$\approx$2-20\,$R_\oplus$.  This size can be converted into a mass
estimate by requiring the optical depth $\tau$ of the cloud to be at
least unity and by positing a cloud composition.  In other
words, a minimum constraint on the number density follows by requiring
the cloud to be optically thick.  For cases like LP 12-502, this is
reasonable because the transit duration of the shortest dips implies $R_{\rm cloud}\ll
R_\star$.  Carrying out the relevant calculation assuming the
occulting material is dust grains 1\,$\mu$m in size,
\citet{2023MNRAS.518.4734S} reported minimum cloud masses of order
$10^{12}$\,kg (their Eq.~23), which scale linearly with both the
optical depth and dust grain radius.  This is comparable to a small
asteroid; the asteroid belt itself has a mass of order
$\approx$10$^{21}$\,kg \citep{2019Icar..319..812P}.  The previously
discussed cool gas prominences would need to have masses of order
$10^{14}$\,kg \citep{1990MNRAS.247..415C}, about 100$\times$ larger
than the requisite dust mass.

Our observations provide a direct measurement of how often dips appear
and disappear, both due to sudden state-switches, and due to more
gradual, secular evolution.  For instance, LP~12-502 showed three
``state-switch'' events over the six months of available TESS
observations, during cycles 248-264, 299-315, and 1233-1247.  In each
case, a dip ``turned off''.  It is plausible to imagine that these
events correspond to either mass being ejected from the corotating
clumps, or else perhaps being accreted onto the star.  In either case,
the corresponding $\dot{M} \equiv M\cdot {\rm d}N/{\rm d}t$
time-averaged over six months is $\approx$3$\times$$10^{-18}
M_\odot\,{\rm yr}^{-1} \approx 1\times10^{-12} M_\oplus\,{\rm
yr}^{-1}$.  Considered cumulatively over the $\approx$$10^8$ years for
which the CQV phenomenon is observed, this yields a cumulative moved
dust mass of $10^{-4}\,M_\oplus$, of order the Solar System's asteroid
belt.  If the occulting material is gas, the masses involved would be
of order 100 times larger.  For cases in which we observe the {\it
growth} of dips, such as the Sector~29 data for TIC~224283342, or
Sector~5 of TIC~294328885, the dip depths typically increase by of
order a few percent over ten to twenty days.  This growth rate yields
a mass flux one order of magnitude larger than the earlier estimate.

@@

An argument against the connection between CQVs and the $\sigma$~Ori~E
analogs is that the light curve of $\sigma$~Ori~E is much simpler than
those in Figure~\ref{fig:cqvs}, with only two broad local minima, and
one ``hump'' (Figure~\ref{fig:bstar}), top-left; see also
\citealt{2022ApJ...924L..10J}).  Within the model proposed by Townsend
et al., this is explained by the simplicity of the star's magnetic
field, which to first-order is a dipole, similar to most other
magnetic B stars \citep{2007A&A...475.1053A,2009ARA&A..47..333D}.
This approximation matches the qualitative behavior of both the
broadband light curves, and the line-profile variations seen in
hydrogen, helium, and carbon \citep{2012MNRAS.419..959O}.

@@
